Comprehensive Quality-Aware Automated Semantic Web Service Composition

### 2017-05-16 paper.tex
update

\textit{Web service composition} aims to loosely couple a set of web services to provide a value-added composite service that accommodates complex functional and non-functional requirements of service users. 
%This application is developed by integrating interoperable and collaborative functionalities over heterogeneous systems. 
Two most notable challenges for web service composition are ensuring interoperability of services and achieving Quality of Service (QoS) optimisation \cite{fensel2011semantic}. \textit{Interoperability} of web services presents challenge in syntactic and semantic dimensions. The syntactic dimension is covered by the XML-based technologies, such as $WSDL$, $SOAP$. The semantic dimension enables a better collaboration through ontology-based semantics, such as OWL-S, WSML, and SAWSDL \cite{petrie2016web}. \textit{Semantic web services composition} is distinguished from the syntactic service composition, as the resources of semantic web services are described semantically to enable a better interoperability for chaining web services. Another challenge is related to QoS optimisation. This problem gives birth to \textit{QoS-aware service composition} that aims to find composition solutions with optimised QoS.

Existing works on service composition focus mainly on addressing only one challenge above. In these works, huge efforts have been devoted to QoS-aware web service compositions assuming a pre-defined abstract workflow is given. This is generally considered as a \textit{semi-automated web service composition} approach. Generating composition plans automatically in discovering and selecting suitable web services is a NP-hard problem \cite{moghaddam2014service}. In the past few years, many approaches \cite{gupta2015optimization,ma2015hybrid,qi2010combining,da2016particle,da2015graphevol,yu2013adaptive} to QoS-aware web service composition employ Evolutionary Computation (EC) techniques to automatically generate composition solutions. Genetic Programming (GP) based approaches produce promising results, but these approaches often require repairing or penalising the solutions \cite{ma2015hybrid,yu2013adaptive}. However, Particle Swarm Optimisation (PSO) is considered to be more efficient, as it simply updates particles' positions to move candidate particles across the searching space, rather than employing genetic operators to carry out searching process. All these works have enabled an \textit{automatic web service composition}, but do not optimise QoS and quality of semantic matchmaking simultaneously to achieve a desirable balance on both sides.

The overall goal of this paper is to \textit{develop a PSO-based approach to comprehensive quality-aware automated semantic web service composition that satisfactorily optimises both QoS and semantic matchmaking quality}. Particularly, this paper extends existing works on QoS-aware service composition by considering jointly optimising the both quality aspects, which is proposed as a comprehensive quality model. We will achieve three objectives in this work:

@@

In this paper we are only interested in robust compositions where only $exact$ and $plugin$ matches are considered, see \cite{lecue2009optimizing}. As argued in \cite{lecue2009optimizing} $plugin$ matches are less preferable than $exact$ matches due to the overheads associated with data processing. We suggest to consider the semantic similarity of concepts when comparing different $plugin$ matches. For concepts $a, b$ in $\mathcal{O}$ the \emph{semantic similarity} $sim(a, b)$ is calculated based on the edge counting method in a taxonomy like WorldNet or Ontology using Eq. (\ref{eq_sim}). This method \cite{shet2012new} has the advantages of simple calculation and good performance.  In Eq. (\ref{eq_sim}), $N_a$, $N_b$ and $N_c$ measure the distances from concept $a$, concept $b$, and the closest common ancestor $c$ of $a$ and $b$ to the top concept of the ontology $\mathcal{O}$, respectively. 
\begin{equation}
sim(a, b){=} \frac{2N_c \cdot e^{-\lambda L/D} }{N_{a}+N_{b}}
\label{eq_sim}
\end{equation}
\noindent For our purposes, $\lambda$ can be set to 0 as we do not measure the similarities of neighbourhood concepts, the matching type not considered in this paper.

@@

\begin{align*}
type_e = 
\begin{cases}
	1 & \text{ if $a\equiv b$ ($exact$ match)}\\
	p & \text{ if $a \sqsubseteq b$ ($plugin$ match)}
\end{cases}
&&
sim_e = sim(a,b) = \frac{2N_c}{N_{a}+N_{b}}
\end{align*}

@@

The \emph{semantic matchmaking quality} of the service composition can be obtained by aggregating $type_e$ and $sim_e$, respectively, over all edges in $E$ as follow:
\begin{align*}
MT {=} \prod_{j=1}^{m} type_ {e_{j}}
&&
SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
\end{align*}

%\begin{equation}
%\label{equation6}
%MT {=} \prod_{j=1}^{m} type_ {e_{j}}
%\end{equation}
%\begin{equation}
%\label{equation7}
%SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
%\end{equation}
The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services \cite{ma2015hybrid}. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliability $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of most time-consuming path in the composition, i.e., $$T = MAX \{\sum\limits^{\ell_n}_{k=1}t_{S_k} | \text{ $\ell_n$ is the length of a path $P_n$ and $n\in\{1,\ldots,n\}$}\}.$$

\noindent When multiple quality criteria are involved into decision making, then the overall fitness of a solution can be defined as a weighted sum of the individual criteria in Eq \ref{eq_fitness}, assuming the preference of each quality criterion is known by users. 
\begin{equation}
\label{eq_fitness}
Fitness = w_1 \hat{MT} + w_2 \hat{SIM} + w_3 \hat{A} + w_4 \hat{R} + w_5(1 - \hat{T}) + w_6(1 - \hat{C})
\end{equation}
\noindent with $\sum_{k=1}^{6} w_k= 1$. We call this objective function the \emph{comprehensive quality model} for service composition.
The weights can be adjusted according to users' preferences. $\hat{MT}$, $\hat{SIM}$, $\hat{A}$, $\hat{R}$, $\hat{T}$, and $\hat{C}$ are normalised values calculated within the range from 0 to 1 using Eq. (\ref{eq_normal}). To simplify the presentation we also use the notation $(Q_1,Q_2,Q_3,Q_4,Q_5,Q_6) = (MT,SIM,A,R,T,C)$. $Q_1$ and $Q_2$ have minimum value 0 and maximum value 1. The minimum and maximum value of $Q_3$, $Q_4$, $Q_5$, and $Q_6$ are calculated across all task-related candidates in the service repository $\mathcal{SR}$ using the greedy search in \cite{ma2015hybrid,da2016genetic}.

@@

The parameters for the PSO are chosen from the settings from \cite{shi2001particle}, In particular, PSO population size is 30 with 100 generations. We run 30 times independently for each dataset. We configure the weights of fitness function to properly balance semantic matchmaking quality and QoS. Therefore, $w_{1}$ and $w_{2}$ are set equally to 0.25, and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125. The $p$ of $type_e$ is set to 0.75 ($plugin$ match) according to \cite{lecue2009optimizing}. In general, weight settings and parameter $p$ are decided by users' preferences.

@@

Recently, a QoS Model, $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$, where $\sum_{i=1}^{4} w_i = 1$, is widely used for QoS-aware web service composition \cite{ma2015hybrid,da2016particle,da2015graphevol}. To show the effectiveness of our proposed comprehensive quality model, we compare the best solutions found by this QoS model and our comprehensive model using our PSO-based approach. We record and compare the mean values of both $SM$ ($SM = 0.5 \hat{MT} + 0.5 \hat{SIM}$) and $QoS$($QoS = 0.25 \hat{A} + 0.25 \hat{R} + 0.25(1 - \hat{T}) + 0.25(1 - \hat{C})$) of best solutions over 30 independent runs. To make the comparison informative, all these recorded values have been normalised from 0 to 1, and compared using independent-samples T tests, see Table \ref{decisionTable}. Note that p-values are smaller than 0.001 in the significance of the results.

@@

To analyse the effectiveness of achieving a good comprehensive quality at the expense of slightly reduced QoS, we demonstrate two best solutions produced using Task 3 as an example. Fig. \ref{comparisontest} $(1)$ and $(2)$ show two weighted DAGs $WG_1$ and $WG_2$ that have been obtained as the best service compositions solutions based on the QoS model and on the comprehensive quality model, respectively. Both $WGs$ have exactly the same service workflow structure, but some service vertices and edges denoted in red are different. To better understand these differences, we list the overall semantic matchmaking quality $SM$,  overall $QoS$ and semantic matchmaking quality $sm_{e_n}$ associated to these different edges in $WG_1$ and $WG_2$. (Note: $sm_{e_n} = 0.5type_{e_n} + 0.5 sim_{e_n}$), where $\Delta Q$ reveals the gain (positive $\Delta Q$) or a loss (negative $\Delta Q$) of the listed qualities for our comprehensive quality model. Therefore, we achieve a comprehensive quality gain (+0.1433), a result of a gain in semantic matchmaking quality (+0.1467) and a loss in $QoS$ (-0.0034). To understand the improvement of semantic matchmaking quality from these numbers, we pick up $e_4$ that is associated with the smallest $\Delta Q$. The $e_4$ of $WG_1$ and $WG_2$ has two different source nodes, $Ser1640238160$ and $Ser947554374$, and two the same end nodes. $Ser1640238160$ and $Ser947554374$ are services with output parameters $Inst582785907$ and  $Inst795998200$ corresponds to two concepts $Con2037585750$ and $Con103314376$ respectively in the given ontology shown in Fig. \ref{comparisontest} $(4)$. As $Inst658772240$ is a required parameter of $End$, and related to concept $Con2113572083$, $Inst795998200$ is closer to the required output $Inst658772240$ than $Inst582785907$. Therefore,  $Ser947554374$ is selected with a better semantic matchmaking quality compared to $Ser1640238160$.

>>>>

\textit{Web service composition} aims to loosely couple a set of web services to provide a value-added composite service that accommodates complex functional and non-functional requirements of service users. 
%This application is developed by integrating interoperable and collaborative functionalities over heterogeneous systems. 
Two most notable challenges for web service composition are ensuring interoperability of services and achieving Quality of Service (QoS) optimisation \cite{fensel2011semantic}. \textit{Interoperability} of web services presents challenge in syntactic and semantic dimensions. The syntactic dimension is covered by the XML-based technologies, such as $WSDL$, $SOAP$. The semantic dimension enables a better collaboration through ontology-based semantics, such as OWL-S, WSML, and SAWSDL \cite{petrie2016web}. \textit{Semantic web services composition} is distinguished from the syntactic service composition, as the resources of semantic web services are described semantically to enable a better interoperability for chaining web services. Another challenge is related to QoS optimisation. This challenge gives birth to \textit{QoS-aware service composition} that aims to find composition solutions with optimised QoS.

Existing works on service composition focus mainly on addressing only one challenge above. In these works, huge efforts have been devoted to QoS-aware web service compositions assuming a pre-defined abstract workflow is given. This is generally considered as a \textit{semi-automated web service composition} approach. Generating composition plans automatically in discovering and selecting suitable web services is a NP-hard problem \cite{moghaddam2014service}. In the past few years, many approaches \cite{gupta2015optimization,ma2015hybrid,qi2010combining,da2016particle,da2015graphevol,yu2013adaptive} to QoS-aware web service composition employ Evolutionary Computation (EC) techniques to automatically generate composition solutions. Genetic Programming (GP) based approaches produce promising results, but these approaches often require repairing or penalising the solutions \cite{ma2015hybrid,yu2013adaptive}, which restrict their searching space. Particle Swarm Optimisation (PSO) based approaches are proposed to avoid producing invalid solutions to effectively search for near-optimal solutions \cite{da2016particle}. All these works have enabled an \textit{automatic web service composition}, but do not optimise QoS and quality of semantic matchmaking simultaneously to achieve a desirable balance on both sides.

The overall goal of this paper is to \textit{develop a PSO-based approach to comprehensive quality-aware automated semantic web service composition that simultaneously optimises both QoS and semantic matchmaking quality}. Particularly, this paper extends existing works on QoS-aware service composition by considering jointly optimising the both quality aspects, which is proposed as a comprehensive quality model. We will achieve three objectives in this work:

@@

In this paper we are only interested in robust compositions where only $exact$ and $plugin$ matches are considered, see \cite{lecue2009optimizing}. As argued in \cite{lecue2009optimizing} $plugin$ matches are less preferable than $exact$ matches due to the overheads associated with data processing. We suggest to consider the semantic similarity of concepts when comparing different $plugin$ matches. For concepts $a, b$ in $\mathcal{O}$ the \emph{semantic similarity} $sim(a, b)$ is calculated based on an edge counting method in a taxonomy like WorldNet or Ontology using Eq. (\ref{eq_sim}) \cite{shet2012new}. This method has the advantages of simple calculation and good performance \cite{shet2012new}. In Eq. (\ref{eq_sim}), $N_a$, $N_b$ and $N_c$ measure the distances from concept $a$, concept $b$, and the closest common ancestor $c$ of $a$ and $b$ to the top concept of the ontology $\mathcal{O}$, respectively. 
\begin{equation}
sim(a, b){=} \frac{2N_c \cdot e^{-\lambda L/D} }{N_{a}+N_{b}}
\label{eq_sim}
\end{equation}
\noindent For our purposes, $\lambda$ can be set to 0 as we do not measure the similarities of neighbourhood concepts, the matching type not considered in this paper.

@@

\begin{align*}
type_e = 
\begin{cases}
	1 & \text{ if $a\equiv b$ ($exact$ match)}\\
	p & \text{ if $a \sqsubseteq b$ ($plugin$ match)}
\end{cases}
,&&
sim_e = sim(a,b) = \frac{2N_c}{N_{a}+N_{b}}
\end{align*}

@@

The \emph{semantic matchmaking quality} of the service composition can be obtained by aggregating $type_e$ and $sim_e$, respectively, over all edges in $E$ as follow:
\begin{align*}
MT {=} \prod_{j=1}^{m} type_ {e_{j}}
,&&
SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
\end{align*}

%\begin{equation}
%\label{equation6}
%MT {=} \prod_{j=1}^{m} type_ {e_{j}}
%\end{equation}
%\begin{equation}
%\label{equation7}
%SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
%\end{equation}
The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services \cite{ma2015hybrid}. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliability $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of the most time-consuming path in the composition, i.e., assuming $h$ is the number of path,
$$T = MAX \{\sum\limits^{\ell_j}_{k=1}t_{S_k} | \text{ $j\in\{1,\ldots,h\}$,  and $\ell_j$ is the length of a path $P_j$}\}$$.
\noindent When multiple quality criteria are involved into decision making, then the overall fitness of a solution can be defined as a weighted sum of the individual criteria in Eq \ref{eq_fitness}, assuming the preference of each quality criterion is provided by users. 
\begin{equation}
\label{eq_fitness}
Fitness = w_1 \hat{MT} + w_2 \hat{SIM} + w_3 \hat{A} + w_4 \hat{R} + w_5(1 - \hat{T}) + w_6(1 - \hat{C})
\end{equation}
\noindent with $\sum_{k=1}^{6} w_k= 1$. We call this objective function the \emph{comprehensive quality model} for service composition.
The weights can be adjusted according to users' preferences. $\hat{MT}$, $\hat{SIM}$, $\hat{A}$, $\hat{R}$, $\hat{T}$, and $\hat{C}$ are normalised values calculated within the range from 0 to 1 using Eq. (\ref{eq_normal}). To simplify the presentation we also use the notation $(Q_1,Q_2,Q_3,Q_4,Q_5,Q_6) = (MT,SIM,A,R,T,C)$. $Q_1$ and $Q_2$ have minimum value 0 and maximum value 1. The minimum and maximum value of $Q_3$, $Q_4$, $Q_5$, and $Q_6$ are calculated across all task-related candidates in the service repository $\mathcal{SR}$ using the greedy search in \cite{ma2015hybrid,da2016genetic}.

@@

The parameters for the PSO are chosen from the settings from \cite{shi2001particle}, In particular, PSO population size is 30 with 100 generations. We run 30 times independently for each dataset. We configure the weights of fitness function to properly balance semantic matchmaking quality and QoS. Therefore, $w_{1}$ and $w_{2}$ are set equally to 0.25, and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125. The $p$ of $type_e$ is set to 0.75 ($plugin$ match) according to \cite{lecue2009optimizing}. In general, weight settings and parameter $p$ are decided according to users' preferences.

@@

Recently, a QoS Model, $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$, where $\sum_{i=1}^{4} w_i = 1$, is widely used for QoS-aware web service composition \cite{ma2015hybrid,da2016particle,da2015graphevol}. To show the effectiveness of our proposed comprehensive quality model, we compare the best solutions found by this QoS model and our comprehensive model using our PSO-based approach. We record and compare the mean values of both $SM$ ($SM = 0.5 \hat{MT} + 0.5 \hat{SIM}$) and $QoS$($QoS = 0.25 \hat{A} + 0.25 \hat{R} + 0.25(1 - \hat{T}) + 0.25(1 - \hat{C})$) of best solutions over 30 independent runs. To make the comparison informative, all these recorded values have been normalised from 0 to 1, and compared using independent-samples T tests, see Table \ref{decisionTable}. Note that p-values are consistently smaller than 0.001 in the results with significant differences.

@@

To analyse the effectiveness of achieving a good comprehensive quality at the expense of slightly reduced QoS, we demonstrate two best solutions produced using Task 3 as an example. Fig. \ref{comparisontest} $(1)$ and $(2)$ show two weighted DAGs, $WG_1$ and $WG_2$, which have been obtained as the best service compositions solutions based on the QoS model and on the comprehensive quality model, respectively. Both $WGs$ have exactly the same service workflow structure, but some service vertices and edges denoted in red are different. To better understand these differences, we list the overall semantic matchmaking quality $SM$,  overall $QoS$ and semantic matchmaking quality $sm_{e_n}$ associated to these different edges in $WG_1$ and $WG_2$. (Note: $sm_{e_n} = 0.5type_{e_n} + 0.5 sim_{e_n}$), where $\Delta Q$ reveals the gain (positive $\Delta Q$) or a loss (negative $\Delta Q$) of the listed qualities for our comprehensive quality model. Therefore, we achieve a comprehensive quality gain (+0.1433), a result of a gain in semantic matchmaking quality (+0.1467) and a loss in $QoS$ (-0.0034). To understand the improvement of semantic matchmaking quality from these numbers, we pick up $e_4$ that is associated with the smallest $\Delta Q$. The $e_4$ of $WG_1$ and $WG_2$ has two different source nodes, $Ser1640238160$ and $Ser947554374$, and two the same end nodes. $Ser1640238160$ and $Ser947554374$ are services with output parameters $Inst582785907$ and  $Inst795998200$ corresponds to two concepts $Con2037585750$ and $Con103314376$ respectively in the given ontology shown in Fig. \ref{comparisontest} $(4)$. As $Inst658772240$ is a required parameter of $End$, and related to concept $Con2113572083$, $Inst795998200$ is closer to the required output $Inst658772240$ than $Inst582785907$. Therefore,  $Ser947554374$ is selected with a better semantic matchmaking quality compared to $Ser1640238160$.
